Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-02-18 draft.tex
Updates from Overleaf

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/SNuM.png}
    \caption{SN Ia rate per unit stellar mass as function of galaxy stellar mass. Error bars are the same as Fig. \ref{fig:rate_raw}. SDSS data (orange squares) are taken from \citet{Graur2013}. The systematically higher SNuM in DES is likely caused by the higher redshift of the sample. }%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:snum}
\end{figure}
By binning both the SN hosts and field galaxies by their stellar mass we measure the mean SN Ia rate per galaxy as a function of stellar mass, $R_{\mathrm{G}}(M_*)$. We employ a bootstrap Monte-Carlo approach in order to estimate the uncertainty in the value of the rate in each stellar mass bin. The probability distribution function (PDF) of each galaxy's stellar mass is represented by the sum of two half Gaussian distributions to represent the asymmetric positive and negative uncertainties derived in the SED fitting stage (Section \ref{sec:data}). For each SN host and field galaxy, we take 100 samples of its stellar mass by drawing at random from its PDF. For each of the 100 samples we calculate the SN rate per galaxy in each stellar mass bin via Eq \ref{eq:rate1}, using the completeness-corrected values of $N_{\mathrm{SN}}$ and $N_{\mathrm{SN}}$ from Eqs. \ref{eq:corr_SN} and \ref{eq:corr_G} respectively.

@@

\textbf{Add plot showing that the toy model recreates the stellar mass function?}

%\textbf{TEMPTED TO DITCH THIS SINCE OUR SFR IS SKETCHY}
%In order to make a more detailed comparison of our results with those of \citet{Sullivan2006} and \citet{Smith2012}, we split galaxies according to their sSFR: galaxies with sSFR$\leq -11.5$ yr$^{-1}$ are denoted as passive, those with $-11.5 < \mathrm{sSFR} \leq -9.5$ are moderately star-forming, and those with sSFR $>-9.5$ are labelled as highly star-forming. The results when split this way are shown in Fig. \ref{fig:rate_raw_split_sfr}. Star-forming galaxies show a steeper slope than their passive counterparts, with values of $1.10 \pm 0.05$ and $0.84 \pm 0.04$ for moderate and highly star-forming galaxies respectively, while passive galaxies show a much shallower slope of $0.55 \pm 0.08$. These values are broadly similar to those measured in SDSS by \citet{Smith2012}, with star-forming galaxies showing steeper slopes than passives (\citet{Smith2012} found 1.01 and 0.72 for SF and passive, respectively). We note that we detect no passive galaxies with mass below $10^{10} \mathrm{M}_{\odot}$. Above this value, the \citet{Smith2012} data for passive galaxies is considerably flatter than that which we measure. On the other hand our results are inconsistent with those found at a similar redshift to the DES data in SLNS \citep{Sullivan2006}, who measured a slope of 1.10 in passives but 0.66 and 0.74 in moderate and highly star-forming galaxies respectively. We note that the moderately star-forming slope is thus consistent between DES and SNLS, but the slopes of passive and highly star-forming galaxies are swapped.
\renewcommand{\arraystretch}{1.2}
\begin{table}
	\centering
	\caption{Results of the Bayesian parameter estimation for the SN Ia DTD.}
	\label{tab:dtd_results}
	\begin{tabular}{lccr} % four columns, alignment for each
		\hline
		 &$\beta$ & $A$ & $t_{\mathrm{p}}$\\
		 &-       & $10^{-13}$ M$_{\odot}^{-1}$yr$^{-1}$ & Gyr \\
		\hline
		Fixed $t_{\mathrm{p}}$ & $-1.13\pm0.5$ &  $2.08 \pm 0.08$ & -\\
		Free $t_{\mathrm{p}}$ & $-1.02 _{-0.11} ^{+0.08}$ & $1.86 _{-0.13} ^{+0.19}$ & $0.067 _{-0.023} ^{+0.043}$\\
		Fixed $A$, $\beta$ & - & - & $0.047_{-0.007}^{+0.008}$\\
		\hline
	\end{tabular}
\end{table}

@@

We assume a DTD $\Phi(\tau)$ with a power-law form described by index $\beta$ (Eq. \ref{eq:dtd}), normalisation $A$ and effective after prompt time $t_{\mathrm{p}}$. As per Section \ref{sec:rates}, we constrain the parameters via Bayesian inference, using HMC. At each step of the sampling procedure the model is re-calculated via Eq. \ref{eq:galaxy_rate}, and evaluate the likelihood assuming that the rate data are normally distributed about their means. For $\beta$ we adopt a Gaussian prior with hyper-parameters $p(\beta) \sim \mathcal{N}(-1\,,0.3)$. We fit for $A$ in log space, and adopt a Gaussian prior (in log space) with hyper-parameters $p(A) \sim \mathcal{N}(-12.7\,, 0.5)$.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/SNuM.png}
    \caption{SN Ia rate per unit stellar mass as function of galaxy stellar mass. Error bars are the same as Fig. \ref{fig:rate_raw}. SDSS data (orange squares) are taken from \citet{Graur2013}. The systematically higher SNuM in DES is caused by the higher redshift of the sample, as illustrated by the DTD model prediction (coloured curves) }%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:snum}
\end{figure}
By binning both the SN hosts and field galaxies by their stellar mass we measure the mean SN Ia rate per galaxy as a function of stellar mass, $R_{\mathrm{G}}(M_*)$. We employ a bootstrap Monte-Carlo approach in order to estimate the uncertainty in the value of the rate in each stellar mass bin. The probability distribution function (PDF) of each galaxy's stellar mass is represented by the sum of two half Gaussian distributions to represent the asymmetric positive and negative uncertainties derived in the SED fitting stage (Section \ref{sec:data}). For each SN host and field galaxy, we take 100 samples of its stellar mass by drawing at random from its PDF. For each of the 100 samples we calculate the SN rate per galaxy in each stellar mass bin via Eq \ref{eq:rate1}, using the completeness-corrected values of $N_{\mathrm{SN}}$ and $N_{\mathrm{SN}}$ from Eqs. \ref{eq:corr_SN} and \ref{eq:corr_G} respectively.

@@

%\textbf{TEMPTED TO DITCH THIS SINCE OUR SFR IS SKETCHY}
%In order to make a more detailed comparison of our results with those of \citet{Sullivan2006} and \citet{Smith2012}, we split galaxies according to their sSFR: galaxies with sSFR$\leq -11.5$ yr$^{-1}$ are denoted as passive, those with $-11.5 < \mathrm{sSFR} \leq -9.5$ are moderately star-forming, and those with sSFR $>-9.5$ are labelled as highly star-forming. The results when split this way are shown in Fig. \ref{fig:rate_raw_split_sfr}. Star-forming galaxies show a steeper slope than their passive counterparts, with values of $1.10 \pm 0.05$ and $0.84 \pm 0.04$ for moderate and highly star-forming galaxies respectively, while passive galaxies show a much shallower slope of $0.55 \pm 0.08$. These values are broadly similar to those measured in SDSS by \citet{Smith2012}, with star-forming galaxies showing steeper slopes than passives (\citet{Smith2012} found 1.01 and 0.72 for SF and passive, respectively). We note that we detect no passive galaxies with mass below $10^{10} \mathrm{M}_{\odot}$. Above this value, the \citet{Smith2012} data for passive galaxies is considerably flatter than that which we measure. On the other hand our results are inconsistent with those found at a similar redshift to the DES data in SLNS \citep{Sullivan2006}, who measured a slope of 1.10 in passives but 0.66 and 0.74 in moderate and highly star-forming galaxies respectively. We note that the moderately star-forming slope is thus consistent between DES and SNLS, but the slopes of passive and highly star-forming galaxies are swapped.
\renewcommand{\arraystretch}{1.2}
\begin{table}
	\centering
	\caption{Results of the Bayesian parameter estimation for the SN Ia DTD.}
	\label{tab:dtd_results}
	\begin{tabular}{lccr} % four columns, alignment for each
		\hline
		 &$\beta$ & $A$ & $t_{\mathrm{p}}$\\
		 &-       & $10^{-13}$ M$_{\odot}^{-1}$yr$^{-1}$ & Gyr \\
		\hline
		Fixed $t_{\mathrm{p}}$ & $-1.13\pm0.5$ &  $2.08 \pm 0.08$ & -\\
		Free $t_{\mathrm{p}}$ & $-1.03 _{-0.07} ^{+0.11}$ & $1.87 _{-0.17} ^{+0.13}$ & $0.07 _{-0.04} ^{+0.03}$\\
		Fixed $A$, $\beta$ & - & - & $0.047_{-0.007}^{+0.008}$\\
		\hline
	\end{tabular}
\end{table}

@@

We assume a DTD $\Phi(\tau)$ with a power-law form described by index $\beta$ (Eq. \ref{eq:dtd}), normalisation $A$ and effective after prompt time $t_{\mathrm{p}}$. As per Section \ref{sec:rates}, we constrain the parameters via Bayesian inference, using HMC to explore the posterior distribution. At each step of the sampling procedure the model is re-calculated via Eq. \ref{eq:galaxy_rate}, we evaluate the likelihood assuming that the rate data are normally distributed about their means. For $\beta$ we adopt a Gaussian prior with hyper-parameters $p(\beta) \sim \mathcal{N}(-1\,,0.3)$. We fit for $A$ in log space, and adopt a Gaussian prior (in log space) with hyper-parameters $p(A) \sim \mathcal{N}(-12.7\,, 0.5)$.
